Orchestrating stencils with the Multi-Stencil Framework

### 2017-03-03 ARTICLES/IJPP/articleR1/conclusion.tex
conclusion

%We have presented the Multi-Stencil Language (MSL), a domain specific language for stencil-based numerical simulations. Compared to other existing stencil DSLs, MSL proposes a higher abstraction level for end-users. MSL offers a way to describe a numerical simulation independently from the type of mesh used. As a result, the language can be used for a larger number of simulations. It helps to clearly separate the description of the simulation from implementation choices, which facilitates reuse of existing languages and libraries, for example languages for data structures optimizations and parallelization strategies.

%In this paper, the formal definition of a multi-stencil program has been given. This formalism helps to understand what is defined into the light grammar of MSL, and also how parallelization can be extracted from it. MSL produces an empty parallel pattern of the numerical simulation described. This pattern indicates where synchronizations are needed, but not how they are performed, which is dependent from the type of mesh and from implementation choices. This second step is left to other existing languages, which are chosen as a back-end of MSL. For example, in this paper the chosen back-end is a code using SkelGIS~\cite{CPE:CPE3494,DBLP:conf/ieeehpcs/HeleneS13,DBLP:conf/europar/CoullonL14}, a templated C++ language for distributed meshes (using MPI), and OpenMP~\cite{660313}.

E%xperiments presented in this paper show that the back-end code produced by MSL, \ie by a mesh-agnostic language, does not introduce overheads up to 16.384 cores, and that the hybrid (data and task) parallelism automatically introduced by the language, which combines SkelGIS and OpenMP as a back-end, improves performance compared to a data parallelization only. Thus, with the same code asked to the user, two different parallel versions of the simulation can be obtained.

%However, if performances are improved by the hybrid parallelization, this performance is also limited by the scheduling strategy used in this paper. Actually, as described in Section~\ref{sect:msp}, a static series-parallel (fork-join) scheduling is performed in this work. This helps to define a predictive performance model, which has been validated in experiments. However, getting rid of global synchronizations (join) could be an interesting improvement to get better performance, especially for irregular simulations, where tasks are not naturally balanced. Our immediate perspective, thus, is to use OpenMP dynamic schedulers~\cite{Gautier:2013:XRS:2510661.2511383,Augonnet2011,wu:hal-01078359} as back-ends to study performance improvments. Using such dynamic schedulers also opens to more heterogeneous architectures, such as GPUs or many-cores accelerators.

%Finally, to validate the separation of concerns introduced by MSL, a second back-end for unstructured meshes could be interesting. We think about PaMPA~\cite{lachat:hal-00768916} as a distributed data structure backend, instead of SkelGIS. Thus using a single language, the MSL language, two different kinds of numerical simulations could be defined: one onto Cartesian meshes; and one on unstructured meshes.

\HC{framework to ease HPC specialist work for proposing new meshes etc.}
\HC{how to choose data or hybrid, change on the fly if needed}
\HC{how to choose if a fusion is good or not, change on the fly or learning strategies onto production codes}
\HC{In addition to this, by using components, doors to online reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} are also open (even if not answered in this paper).}

>>>>

In this paper, we have presented the Multi-Stencil Framework. MSF is built upon a meta-model of a multi-stencil program which has been presented in Section~\ref{sect:formalism}. From this meta-model has been designed, first, the generic component assembly of a multi-stencil program, and second, the domain specific language MSL that enables the description of a specific application by a numerician. From these entries, the MSF compiler, namely MSC, is able to generate automatically a parallel specific assembly of the simulation. This specific assembly represents the parallel orchestration of computations independently of implementation choices. Moreover, by combining a meta-model and component-based programming models MSF has the particularity to enhance separation of concerns and code-reuse of existing solutions (\eg SkelGIS and OpenMP in this paper). Finally, MSF has been evaluated on the real case simulation FullSWOF2D. Results have shown that even if component models have been used, strong and weak scalings of MSF (for its data parallelization technique) do not suffer from overheads. Results have also shown that the hybrid parallelization handled by MSF could increase performances when MPI communications are too heavy, which is new compared to existing solutions. Finally, we have presented evaluations on fusion optimization that demonstrates that fusion is not always a good choice, particularly when performing hybrid parallelism.

Many perspectives of future work raise from this work. First, even if using one distributed data structure or another is possible and facilitated by MSF, the HPC specialist still has to understand component programming models to add one. We think it could be interesting to also facilitate the work of the HPC specialist by proposing a semi-automated framework to add external tools to MSF.

Second, it has been shown in the performance model and evaluation results that the choice between data parallelism and hybrid parallelism in not trivial. It depends on the time spent in computations, the number of cores and nodes used for the execution, as well as the network behavior (Fig.~\ref{fig:limit} and~\ref{fig:close}). MSF, thanks to its meta-model and to the use of component models, handles the activation of one parallelism technique or another. Thus MSF can be considered as a performance leverage for multi-stencil programs. However, it would be interesting to investigate how an actor algorithm could choose the best parallelization technique for a given execution of a production multi-stencil simulation.

The same choice problem is raised by fusion evaluations of Section~\ref{sect:fus}. Actually, fusion optimization is not always relevant. Thus, for the same reasons than parallelization techniques, an actor algorithm could choose the best solution for a given execution. To design such actor algorithms different solutions should be studied, such as following an approximated behavior model, or using setup runs to use deep learning algorithms etc.

Finally, by using component models doors are opened to reconfiguration of applications~\cite{Lanore:2015:RCM:2737166.2737169} which means that the application could adapt its behavior and structure to external events.
